Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-05-28 thesis/chapters/ch_introduction.tex
started writing appendxi to intro to mention other work (cluster Lyalpha, RPS / feedback in dwarfs, grackle) but decided to not do this
Edits from KVJ, JT, and MM for abstract and intro

\chapter[Introduction]{Introduction}
\label{ch:intro}
\vspace{-16pt}
\begin{chapquote}{Isaac Asimov} \singlespacing The most exciting phrase to hear in science, the one that heralds discoveries, is not ‘Eureka!’ but ‘Now that’s funny…
\end{chapquote} \vspace{-8pt}
\begin{chapquote}{Brian W. Kernighan} \singlespacing Debugging is twice as hard as writing the code in the first place. Therefore, if you write the code as cleverly as possible, you are, by definition, not smart enough to debug it
\end{chapquote} \vspace{-8pt}

@@

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, all of the elements in the Universe are produced in nuclear reactions associated with stellar evolution, either in the cores of stars, stellar winds, asymptotic giant branch (AGB) winds of low mass stars, and more exotic sites like neutron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012,AndrewsMartini2013}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{Searle1971,Shaver1983,Belfiore2017,SanchezMenguiano2017}, and detailed stellar abundances in nearby dwarf galaxies with spectroscopically resolved stars \citep[see the review in][]{Tolstoy2009}. Yet in spite of increased number and quality of observations tracing galactic chemical evolution, developing a complete model of galactic chemical evolution is still a daunting task.

%Developing a complete model for galactic chemical evolution is a daunting task.
To give an impression of the difficulties involved, we outline a recipe for constructing a complete model of galactic chemical evolution. One must first prescribe a galaxy's connection to its large scale structure to understand the inflow of pristine, metal-free gas and growth of its gaseous disk, and the accretion (via mergers) of stars formed previously in galaxies hosted by different dark matter halos. One can then worry about knowing where, when, how, and with what masses stars form in the first place from this gas. This requires understanding the hydrodynamic
%and magnetohydrodynamic
properties of the multi-scale, multi-phase turbulent interstellar medium (ISM) and the radiative processes and chemistry that operate within the ISM. This in turn requires one to toss in a detailed understanding of the stellar feedback physics -- stellar radiation, stellar winds, and supernovae -- that helps to regulate star formation by destroying cold, star forming gas, driving the multi-phase structure of the ISM, and by driving outflows of gas into the circumgalactic medium (CGM) around galaxies. Simultaneously, one must make the simple step of nailing down a complete model of stellar structure, stellar evolution, the reaction rates and cross sections of every nuclear reaction, and the lifetimes of individual isotopes. Once all of this is understood, one can then color in the gas in the galaxy over time with the individual metal yields of stellar populations released over their lifetime. With a complete understanding of stellar feedback and the ISM, one can then be sure that they completely capture the mixing of these metals in the ISM over time, and will produce a stellar population with accurate metal abundance ratios. In addition, one would then produce realistic galactic winds, removing the correct amount of metals from the galaxy and enriching the CGM and the intergalactic medium (IGM). Finally, one can layer this model ad nauseum on top of a $\Lambda$CDM model for the cosmological evolution of many galaxies, reproducing all observed properties of galactic chemical evolution as a function of both mass and redshift..... Wait... I forgot about magnetic fields, active galactic nuclei, and cosmic rays...

@@

The observational landscape today is primed for developing a more detailed understanding these processes. Recent observational campaigns such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, the Gaia-ESO survey \citep{Gaia}, APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, GALAH \citep{GALAH,Buder2018}, as well as upcoming observations, such as the Local Volume Mapper as part of SDSS-V, have generated tremendous amounts of information on detailed stellar abundances and stellar kinematics in our Milky Way and nearby Local Group dwarf galaxies. One of the most powerful proposed uses of this enormous trove of data is chemical tagging \citep{Freeman2002}, whereby stellar populations are analyzed in chemical space and 6D phase space to identify co-eval and co-natal groups of stars. This process of galactic archeology aims to break down and identify each distinct stellar component of our Galaxy, explaining the process of their formation and evolution. Substantial work has been made recently to determine the efficacy of this approach \citep[e.g.][]{Ting2012,Hogg2016,Jofre2017,Price-Jones2018,Armillotta2018}, yet the physical processes that give rise to stellar abundances as we observe them today are still uncertain.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.75\linewidth]{./figures/intro/periodic_table}
  \caption{The periodic table of elements colored by the relative fraction of the nucleosynthetic processes that naturally occur in the Universe that produce each element. \textbf{Image Credit:} A modified version of the table created by Anna Frebel (\url{https://www.dropbox.com/sh/ib3agt2bkxqqk2u/AACbEw1warp74qzUh2AcWqfja?dl=0}), accessed on 03-18-19.}
  \label{intro:fig:periodic table}
\end{figure}

A variety of nucleosynthetic processes occur within stars and during their deaths that lead to the production of all of the elements in the periodic table. This is illustrated well in Figure~\ref{intro:fig:periodic table}, which shows the periodic table colored by the fractional importance of the nucleosynthetic processes that produce each element. Due to the diversity in production sources across the periodic table, multi-element metal abundance measurements in stars can be used to to better understand and constrain how and where each metal is produced. Exactly which reactions occur where depend strongly on the initial abundances of elements in stars, their stellar masses, and, for heavy nuclei, the neutron density.

This means that elements from different nucleosynthetic channels are produced on different timescales during a galaxy's evolution, as set by the lifetimes of the stars in which they occur, and are released into the galaxy at different energies, with different spatial distributions within the galaxy, and at different rates. For example, O, an $\alpha$-element is produced predominately in massive stars and is therefore released via core collapse supernovae on timescales of $\sim$~10~Myr after a star formation event, near their stellar birth sites. By contrast, Sr is produced predominately via the s-process (discussed more below) which occurs in part in the asymptotic giant branch (AGB) winds of lower mass stars ($M < 8$~M$_{\odot}$) at much lower energies and on timescales of $\sim 100$~Myr - $1$~Gyr. By this time, these stars would have diffused farther away from their birth sites, potentially sampling a qualitatively different portion of the ISM than the preceeding core collapse SNe. Generally, how different timescales affect observed stellar abundances patterns are well understood and easily modeled.\footnote{The canonical example of this is the ``knee" feature in a $[\alpha/$Fe] vs. [Fe/H] diagram, where the location of this knee signifies the point at which Type Ia enrichment (which produces significant Fe, but little to no $\alpha$ elements) begins to have a significant contribution over core collapse SNe (rich in $\alpha$ elements, but not Fe) \citep[e.g.[]{MatteucciBrocato1990, Geisler2005, Hill2018}. The location of this knee in this diagram shifts for a given galaxy depending on its SFH.} However, it is not well understood how differences in where and with what energies these events occur affect observed stellar abundance patterns. Investigating these differences is a key motivation of this \Dissertation.

As relevant to this work, we briefly summarize a few of the important nucleosynthetic channels shown in Figure~\ref{intro:fig:periodic table} and their sites. $\alpha$-elements, notably O, Mg, Si, Ca, and Ti, are produced during the lifetime of massive stars (O, Mg) or during the core collapse supernova explosion itself. In either case, these elements are synthesized through the capture of an $\alpha$ nucleus, progressing towards increasingly heavier nuclei in intervals of 4/2 in mass/atomic number (as $\alpha$ = $^{4}_2$He) from $^8_4$Be all the way to the stable $^{52}_{26}$Fe and unstable $^{56}_{28}$Ni (which decays to $^{56}_{26}$Fe). The iron group elements, notably Fe, Mn, Co, and Ni, are produced predominately in Type Ia supernovae, with variations depending on a single vs. double degenerate scenario, the initial composition of exploding white dwarf(s) (WD), and the dynamics of the explosion itself.

The presence of these heavier elements in stars enables the formation of even heavier nuclei through neutron capture. This is categorized into two processes, the slow (or s-) and rapid (or r-) process. In the former, neutron densities are typically low enough that the product of each neutron capture beta decays to a more stable isotope before capturing another nucleon. In r-process, however, the rate of neutron capture exceeds beta decay, allowing for the production of different, typically heavier, nuclei. As shown in the diagram, lighter elements are produced via the s-process and heavier via the r-process, with notable overlap in certain elements, such as Ba.\footnote{For the sake of this discussion, the light neutron-capture primary process elements (sometimes referred to as ``lighter element primary process") should be considered s-process elements. Briefly, this different nucleosynthetic channel, first identified in \cite{Travaglio2004}, arose from the disagreement between models of s-process enrichment and stellar abundances of elements including Sr, Y, and Zr. However there is still uncertainty as to whether or not these differences can be explained with different models of weak s-process (which can occur in the winds of massive stars) \citep[e.g.][]{Prantzos2018}.} The s-process occurs predominately in the low-energy winds of low mass (1-8 M$_{\odot}$) AGB stars on time scales of 100's of Myr up to 1 Gyr. Exactly which elements are produced in a given AGB star varies with stellar mass and depends strongly on metallicity, which determines the available heavy seed nuclei. Sr, Y, Zr, and Ba, as well as C, N and F, are commonly used as observational tracers of this nucleosynthetic channel.

The dominant origin of r-process enrichment is highly uncertain, though possible channels include NS-NS and neutron star - black hole mergers, neutrino driven winds in core collapse SNe, and exotic SNe, such as magnetic, jet-driven SNe, collapsars, hypernovae, and long-duration gamma-ray bursts \citep[see ][ for recent reviews]{Frebel2018, Cowan2019}. Eu is the most commonly used (and easy to observe) unambiguous observational tracer of r-process enrichment; Ba is also often studied as a tracer, but requires care due to the significant contribution from s-process enrichment. More detailed spectra \citep[e.g.][]{Ji2018a} can trace many more of these elements, which can be used as an important discriminator between nucleosynthetic models.

@@

A full, detailed model of radiative cooling, heating, non-equillibrium chemistry, and dust is generally far too computationally expensive to include in galaxy-scale simulations. For this reason, many large-scale cosmological simulations assume that all gas is optically thin and that all molecules and ionization states exist in equilibrium. This allows for the quick use of tabulated look-up tables for radiative cooling and heating, often produced using \code{cloudy} \citep{Cloudy2013}, that depend on gas density, temperature, metallicity, and redshift (which sets the UVB heating rate). More complex simulations often track some number (usually $< 10$) of species involved in primordial chemistry in order to better account for non-equilibrium effects and \Hmolecular~production, though with noted increases in memory requirements and computational cost. Some of the most advance models to-date can account for well over 10$^2$ individual chemical reactions involving $>$ 20 elements and molecules \citep[e.g.][]{Richings2016,Richings2018}; but these are typically prohibitively expensive to run on large scales. In this \dissertation we adopt the chemistry and radiative cooling and heating physics models available in the newly developed, open-source library \code{GRACKLE} \citep{GrackleMethod}. The exact physics included in our simulations are discussed in greater detail in Chapter~\ref{ch:chapter1}, but we note that improvements to \code{GRACKLE} have been made as the direct result of the research conducted during this \Dissertation.
%mention that I made improvements to this?

@@

In galaxy-scale hydrodynamics, the costs in resolving the spatial scales and gas densities required to directly follow star formation relegates this process to a sub-grid model. Stars form from fluid elements that surpass a variety of threshold conditions at an assumed rate and efficiency. Usually, simulators set these thresholds to fully encompass gas whose Jeans-length (and thus dynamical behavior) becomes unresolved according to the widely-used \cite{Truelove1997} criterion. Often, gas is required to surpass a simple mass or number density threshold, on top of which simulators can also require that the local gas cells be in a converging flow ($\nabla \cdot \bm{v} < 0$), that gas surpass some virial threshold (describing how bound a cloud may be)
%cite?,
or that gas meet a minimum \Hmolecular~mass fraction \citep[e.g.][]{Kuhlen2012}
%\citep[e.g.][]{}
. Once a fluid element is flagged as star forming, simulators have a few options to choose from in how to determine what fraction and with what rate that mass is converted into stars. Briefly, some form of the following equation is usually adopted to describe the star formation rate density,
\begin{equation}
\label{intro:eq:sfr}
 \dot{\rho}_{\rm SF} = \epsilon \frac{\rho_{\rm gas}}{t_*},
\end{equation}
where $\rho_{\rm gas}$ is the local gas density, $\epsilon$ is some fixed or variable star formation efficiency, and the timescale, $t_*$, is often adopted as the local free fall time, or
\begin{equation}
  t_* = t_{\rm ff} = \sqrt{\frac{3\pi}{32 \rm{G} \rho_{\rm gas}}}.
\end{equation}
In many cases, $\epsilon$ is fixed to some small value ($\sim$ 0.01) \citep{KrumholzTan2007} meant to represent a galaxy-scale average of star formation efficiency, even though this may vary significantly between individual molecular clouds \citep{Grudic2018}. High-resolution simulations, where the mass of stars that would form in a single time step ($dt \times \rm{SFR}$) is far less than a single star particle mass, allow stars to form out of star forming fluid elements stochastically \citep[e.g.][]{Goldbaum2015}.

@@

Initial models for stellar feedback relied primarily on the energy injection from supernovae (SNe, see Section~\ref{intro:sec:supernovae}) to regulate star formation and drive outflows, but it has become clear that this mechanism alone cannot completely account for all of the star formation, ISM, outflow, CGM, and even intergalactic medium (IGM) properties of observed galaxies (see \cite{SomervilleDave2015} and \cite{NaabOstriker2017} for recent reviews and more detailed discussion of how feedback impacts outstanding problems in galactic evolution). Other sources of effective stellar feedback have been identified since \citep[e.g.][]{Agertz2013}, which serve to regulate star formation on different timescales and help develop and ISM and CGM with different phase and ionization properties. This includes both stellar radiation (Section~\ref{intro:sec:radiation}) and stellar winds (Section~\ref{intro:sec:stellarwinds}).

Although not included in this \Dissertation, for the sake of completeness, we mention a few additional sources of stellar feedback that likely contribute to the full picture of galactic evolution. A substantial amount of recent research has been dedicated towards the examination of supernova-driven cosmic ray feedback (see Section~\ref{ch1:sec:CRs} for a more detailed discussion and references). In brief, cosmic rays can be a dominant source of heating in very dense, optically-thick gas and act as a source of non-thermal pressure in the ISM, which can drive galactic winds with qualitatively different phase structures \citep[e.g.][]{SalemBryanCorlies}. Binarity is often ignored in models of stellar feedback. The effects of binary evolution can dramatically extend the timescales over which ionizing radiation and supernovae act in a newly formed stellar populations. Typically, this increases their total energy output and decreases the ISM densities in which SNe explode, increasing their effectiveness (see Section~\ref{ch1:sec:binary stars}). Finally, exotic sources of stellar feedback, such as hypernovae, high-mass X-ray binaries, and NS-NS mergers can also be a potentially important sources of feedback, particularly in the lowest mass halos \citep[e.g.][]{Artale2015}. However, the uncertainties in the frequency and delay time distributions for when these events should occur, and their relative rarity, preclude including these effects in this \Dissertation.

\subsubsection{Supernova Feedback}
\label{intro:sec:supernovae}

In the simplest model, supernova feedback is treated as the injection of pure thermal energy (and mass) into the fluid element(s) that are host / closest to the chosen site of a supernova explosion. In self-consistent models of star formation and feedback these sites are usually actual star particles, but this does not have to be the case. The first models which included supernova feedback lacked the spatial / mass resolution to resolve individual (or even collective) SN events. This is the source of the commonly known ``overcooling" problem, whereby too little energy is injected into too much mass such that the affected region only reaches modest temperatures ($T \lesssim 10^5$~K, as opposed to $T > 10^7$~K, or so). This temperature is right around the peak of the cooling curve, and thus the energy is rapidly radiated away before it has time to make a significant dynamical impact of the galaxy. This is still a problem today in large-box simulations of cosmological volumes which have insufficient resolution to resolve individual SNe. While some of the initial, ad-hoc fixes for this problem (e.g. turning off cooling for some time in SN affected fluid elements or decoupling SN affected SPH particles from the hydrodynamics for some time) are still in use today, some notable advancements have been made to employ some combination of kinetic, momentum, and thermal energy feedback to attempt to capture SNe in a more consistent fashion \citep[e.g.][]{FIRE,Simpson2015,Hopkins2018}.

High resolution simulations, like those developed in this \Dissertation, can typically rely on using pure thermal energy injection while still avoiding artificial overcooling.\footnote{This is advantageous because it is often substantially easier to implement than kinetic / momentum injection methods while being less susceptible to numerical artifacts, especially for grid-based codes.} This is true for simulations with typical spatial / mass resolutions better than a few pc / 10 M$_{\odot}$ \citep{Simpson2015,Hu2018,Smith2018b}. However, a spatial resolution requirement is a density dependent statement; in this case supernovae that occur in the densest gas ($n > 10^{2-3}$~cm$^{-3}$) may still be under-resolved and ineffective. In this case, accounting for feedback processes from massive stars that occur before the first core collapse SN in a given star formation event becomes critical \citep{Hu2016}. Including these physics, as discussed below, greatly reduces the typical ISM densities in which SNe occur, greatly increasing the likelihood that these events are well resolved.

@@

>>>>

\chapter[Introduction]{Introduction}
\label{ch:intro}
\vspace{-16pt}
\begin{chapquote}{Isaac Asimov} \singlespacing The most exciting phrase to hear in science, the one that heralds discoveries, is not ``Eureka!'' but ``Now that's'' funny…
\end{chapquote} \vspace{-8pt}
\begin{chapquote}{Brian W. Kernighan} \singlespacing Debugging is twice as hard as writing the code in the first place. Therefore, if you write the code as cleverly as possible, you are, by definition, not smart enough to debug it
\end{chapquote} \vspace{-8pt}

@@

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, all of the elements in the Universe are produced in nuclear reactions associated with stellar evolution -- in the cores of stars -- and released in stellar winds, asymptotic giant branch (AGB) winds of low mass stars, SNe, and more exotic sources, like neutron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012,AndrewsMartini2013}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{Searle1971,Shaver1983,Belfiore2017,SanchezMenguiano2017}, and detailed stellar abundances in nearby dwarf galaxies with spectroscopically resolved stars \citep[see the review in][]{Tolstoy2009}. Yet in spite of increased number and quality of observations tracing galactic chemical evolution, developing a complete model of galactic chemical evolution is still a daunting task.

%Developing a complete model for galactic chemical evolution is a daunting task.
To give an impression of the difficulties involved, we outline a recipe for constructing a complete model of galactic chemical evolution. One must first prescribe a galaxy's connection to its large scale structure to understand the inflow of pristine, metal-free gas and growth of its gaseous disk, and the accretion (via mergers) of stars formed previously in galaxies hosted by different dark matter halos. One can then worry about knowing where, when, how, and with what masses stars form in the first place from this gas. This requires understanding the hydrodynamic
%and magnetohydrodynamic
properties of the multi-scale, multi-phase turbulent interstellar medium (ISM) and the radiative processes and chemistry that operate within the ISM. This in turn requires one to toss in a detailed understanding of the stellar feedback physics -- stellar radiation, stellar winds, and SNe -- that helps to regulate star formation by destroying cold, star forming gas, driving the multi-phase structure of the ISM, and by driving outflows of gas into the circumgalactic medium (CGM) around galaxies. Simultaneously, one must make the simple step of nailing down a complete model of stellar structure, stellar evolution, the reaction rates and cross sections of every nuclear reaction, and the lifetimes of individual isotopes. Once all of this is understood, one can then color in the gas in the galaxy over time with the individual metal yields of stellar populations released over their lifetime. With a complete understanding of stellar feedback and the ISM, one can then be sure that they completely capture the mixing of these metals in the ISM over time, and will produce a stellar population with accurate metal abundance ratios. In addition, one would then produce realistic galactic winds, removing the correct amount of metals from the galaxy and enriching the CGM and the intergalactic medium (IGM). Finally, one can layer this model ad nauseum on top of a $\Lambda$CDM model for the cosmological evolution of many galaxies, reproducing all observed properties of galactic chemical evolution as a function of both mass and redshift..... Wait... I forgot about magnetic fields, active galactic nuclei, and cosmic rays...

@@

The observational landscape today is primed for developing a more detailed understanding of these processes. Recent observational campaigns such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, the Gaia-ESO survey \citep{Gaia}, APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, GALAH \citep{GALAH,Buder2018}, as well as upcoming observations, such as the Local Volume Mapper as part of SDSS-V, have generated tremendous amounts of information on detailed stellar abundances and stellar kinematics in our Milky Way and nearby Local Group dwarf galaxies. One of the most powerful proposed uses of this enormous trove of data is chemical tagging \citep{Freeman2002}, whereby stellar populations are analyzed in chemical space and 6D phase space to identify co-eval and co-natal groups of stars. This process of galactic archeology aims to break down and identify each distinct stellar component of our Galaxy, explaining the process of their formation and evolution. Substantial work has been made recently to determine the efficacy of this approach \citep[e.g.][]{Ting2012,Hogg2016,Jofre2017,Price-Jones2018,Armillotta2018}, yet the physical processes that give rise to stellar abundances as we observe them today are still uncertain.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.75\linewidth]{./figures/intro/periodic_table}
  \caption{The periodic table of elements colored by the relative fraction of the astrophysical source for each element. \textbf{Image Credit:} Figure 1 of \cite{Johnson2019}.}
  \label{intro:fig:periodic table}
\end{figure}

A variety of nucleosynthetic processes occur within stars and during their deaths that lead to the production of all of the elements in the periodic table. Different astrophysical sources reach conditions that enable different nuclear reactions, as is well-illustrated in Figure~\ref{intro:fig:periodic table}. This image shows the periodic table colored by the fractional importance of the astrophysical source that produce each element. Due to the diversity in production sources across the periodic table, multi-element metal abundance measurements in stars can be used to better understand and constrain how and where each metal is produced. Exactly which reactions occur where depend strongly on the initial abundances of elements in stars, their stellar masses, and, for heavy nuclei, the neutron density.

This means that elements from different nucleosynthetic channels are produced on different timescales during a galaxy's evolution, as set by the lifetimes of the stars in which they occur, and are released into the galaxy at different energies, with different spatial distributions within the galaxy, and at different rates. For example, O, an $\alpha$-element is produced predominantly in massive stars and is therefore released via core collapse SNe on timescales of $\sim$~10~Myr after a star formation event, near their stellar birth sites. By contrast, Sr is produced predominately via the s-process (discussed more below) which occurs in low mass stars and released in asymptotic giant branch (AGB) winds ($M < 8$~M$_{\odot}$) at much lower energies and on timescales of $\sim 100$~Myr - $1$~Gyr. By this time, these stars would have diffused farther away from their birth sites, potentially sampling a qualitatively different portion of the ISM than the preceding core collapse SNe. Generally, how different timescales affect observed stellar abundances patterns are well understood and easily modeled.\footnote{The canonical example of this is the ``knee" feature in a $[\alpha/$Fe] vs. [Fe/H] diagram, where the location of this knee signifies the point at which Type Ia enrichment (which produces significant Fe, but little to no $\alpha$ elements) begins to have a significant contribution over core collapse SNe (rich in $\alpha$ elements, but not Fe) \citep[e.g.][]{MatteucciBrocato1990, Geisler2005, Hill2018}. The location of this knee in this diagram shifts for a given galaxy depending on its SFH.} However, it is not well understood how differences in where and with what energies these events occur affect observed stellar abundance patterns. Investigating these differences is a key motivation of this \Dissertation.

As relevant to this work, we briefly summarize a few of the important nucleosynthetic channels shown in Figure~\ref{intro:fig:periodic table} and their sites. $\alpha$-elements, notably O, Mg, Si, Ca, and Ti, are produced during the lifetime of massive stars (O, Mg) or during the core collapse SN explosion itself. In either case, these elements are synthesized through the capture of an $\alpha$ nucleus, progressing towards increasingly heavier nuclei in intervals of 4/2 in mass/atomic number (as $\alpha$ = $^{4}_2$He) from $^8_4$Be all the way to the stable $^{52}_{26}$Fe and unstable $^{56}_{28}$Ni (which decays to $^{56}_{26}$Fe). The iron group elements, notably Fe, Mn, Co, and Ni, are produced predominately in Type Ia SNe, with variations depending on a single vs. double degenerate scenario, the initial composition of exploding white dwarf(s) (WD), and the dynamics of the explosion itself.

The presence of these heavier elements in stars enables the formation of even heavier nuclei through neutron capture. This is categorized into two processes, the slow (or s-) and rapid (or r-) process. In the former, neutron densities are typically low enough that the product of each neutron capture beta decays to a more stable isotope before capturing another nucleon. In r-process, however, the rate of neutron capture exceeds beta decay, allowing for the production of different, typically heavier, nuclei. The lighter heavy nuclei elements (Rb through Pb, though with exceptions) are produced via the s-process and heavier (Te through Bi, again, with exceptions) via the r-process, with notable overlap in certain elements, such as Ba.
%\footnote{For the sake of this discussion, the light neutron-capture primary process elements (sometimes referred to as ``lighter element primary process") should be considered s-process elements. Briefly, this different nucleosynthetic channel, first identified in \cite{Travaglio2004}, arose from the disagreement between models of s-process enrichment and stellar abundances of elements including Sr, Y, and Zr. However there is still uncertainty as to whether or not these differences can be explained with different models of weak s-process (which can occur in the winds of massive stars) \citep[e.g.][]{Prantzos2018}.}
The s-process occurs predominately in low mass (1-8 M$_{\odot}$) stars and is released at the end of their lives via AGB winds on time scales of 100's of Myr up to 1 Gyr. Exactly which elements are produced in a given AGB star varies with stellar mass and depends strongly on metallicity, which determines the available heavy seed nuclei. Sr, Y, Zr, and Ba, as well as C, N and F, are commonly used as observational tracers of this nucleosynthetic channel.

The dominant origin of r-process enrichment is highly uncertain, though (as indicated in Figure~\ref{fig:periodic table}) possible channels include NS-NS mergers, neutron star - black hole mergers, neutrino driven winds in core collapse SNe, and exotic SNe, such as magnetic, jet-driven SNe, collapsars, hypernovae, and long-duration gamma-ray bursts \citep[see ][ for recent reviews]{Frebel2018, Cowan2019}. Eu is the most commonly used (and easy to observe) unambiguous observational tracer of r-process enrichment; Ba is also often studied as a tracer, but requires care due to the significant contribution from s-process enrichment. More detailed spectra \citep[e.g.][]{Ji2018a} can trace many more of these elements, which can be used as an important discriminator between nucleosynthetic models.

@@

A full, detailed model of radiative cooling, heating, non-equilibrium chemistry, and dust is generally far too computationally expensive to include in galaxy-scale simulations. For this reason, many large-scale cosmological simulations assume that all gas is optically thin and that all molecules and ionization states exist in equilibrium. This allows for the quick use of tabulated look-up tables for radiative cooling and heating, often produced using \code{cloudy} \citep{Cloudy2013}, that depend on gas density, temperature, metallicity, and redshift (which sets the UVB heating rate). More complex simulations often track some number (usually $< 10$) of species involved in primordial chemistry in order to better account for non-equilibrium effects and \Hmolecular~production, though with noted increases in memory requirements and computational cost. Some of the most advanced models to-date can account for well over 10$^2$ individual chemical reactions involving $>$ 20 elements and molecules \citep[e.g.][]{Richings2016,Richings2018}; but these are typically prohibitively expensive to run on large scales. In this \dissertation we adopt the chemistry and radiative cooling and heating physics models available in the newly developed, open-source library \code{GRACKLE} \citep{GrackleMethod}. The exact physics included in our simulations are discussed in greater detail in Chapter~\ref{ch:chapter1}, but we note that improvements to \code{GRACKLE} have been made as the direct result of the research conducted during this \Dissertation.
%mention that I made improvements to this?

@@

In galaxy-scale hydrodynamics, the extreme computational cost of resolving the spatial scales and gas densities required to directly follow star formation relegates this process to a sub-grid model. Stars form from fluid elements that surpass a variety of threshold conditions at an assumed rate and efficiency. Usually, simulators set these thresholds to fully encompass gas whose Jeans-length (and thus dynamical behavior) becomes unresolved according to the widely-used \cite{Truelove1997} criterion. Often, gas is required to surpass a simple mass or number density threshold, on top of which simulators can also require that the local gas cells be in a converging flow ($\nabla \cdot \bm{v} < 0$), that gas surpass some virial threshold (describing how bound a cloud may be)
%cite?,
or that gas meet a minimum \Hmolecular~mass fraction or that density high enough to support \Hmolecular~formation \citep[e.g.][]{Kuhlen2012}.
%\citep[e.g.][]{}
Once a fluid element is flagged as star forming, simulators have a few options for how to compute the fraction and rate at which gas is turned into stars. Briefly, some form of the following equation is usually adopted to describe the star formation rate density,
\begin{equation}
\label{intro:eq:sfr}
 \dot{\rho}_{\rm SF} = \epsilon \frac{\rho_{\rm gas}}{t_*},
\end{equation}
where $\rho_{\rm gas}$ is the local gas density, $\epsilon$ is some fixed or variable star formation efficiency, and the timescale, $t_*$, is often adopted as the local free fall time, or
\begin{equation}
  t_* = t_{\rm ff} = \sqrt{\frac{3\pi}{32 \rm{G} \rho_{\rm gas}}}.
\end{equation}
% AE still need to address mm comments here :
In many cases, $\epsilon$ is fixed to some small value ($\sim$ 0.01) \citep{KrumholzTan2007} meant to represent a galaxy-scale average of star formation efficiency, even though this may vary significantly between individual molecular clouds \citep{Grudic2018}. High-resolution simulations, where the mass of stars that would form in a single time step ($dt \times \rm{SFR}$) is far less than a single star particle mass, allow stars to form out of star forming fluid elements stochastically \citep[e.g.][]{Goldbaum2015}.

@@

Initial models for stellar feedback relied primarily on the energy injection from supernovae (SNe, see Section~\ref{intro:sec:SNe}) to regulate star formation and drive outflows, but it has become clear that this mechanism alone cannot completely account for all of the star formation, ISM, outflow, CGM, and even intergalactic medium (IGM) properties of observed galaxies (see \cite{SomervilleDave2015} and \cite{NaabOstriker2017} for recent reviews and more detailed discussion of how feedback impacts outstanding problems in galactic evolution). Other sources of effective stellar feedback have been identified \citep[e.g.][]{Agertz2013}, which serve to regulate star formation on different timescales and help develop an ISM and CGM with different phase and ionization properties. This includes both stellar radiation (Section~\ref{intro:sec:radiation}) and stellar winds (Section~\ref{intro:sec:stellarwinds}).

Although not included in this \Dissertation, for the sake of completeness, we mention a few additional sources of stellar feedback that likely contribute to the full picture of galactic evolution: 1) cosmic rays, 2) stellar binary evolution, and 3) hypernovae and other exotic sources of feedback. First, a substantial amount of recent research has been dedicated to the examination of SN-driven cosmic ray feedback (see Section~\ref{ch1:sec:CRs} for a more detailed discussion and references). In brief, cosmic rays can be a dominant source of heating in very dense, optically-thick gas and act as a source of non-thermal pressure in the ISM, which can drive galactic winds with qualitatively different phase structures \citep[e.g.][]{SalemBryanCorlies}. Second, binarity is often ignored in models of stellar feedback. The effects of binary evolution can dramatically extend the timescales over which ionizing radiation and SNe act in a newly formed stellar population. Typically, this increases their total energy output and decreases the ISM densities in which SN explode, increasing their effectiveness (see Section~\ref{ch1:sec:binary stars}). Third, exotic sources of stellar feedback, such as hypernovae, high-mass X-ray binaries, and NS-NS mergers can also be a potentially important sources of feedback, particularly in the lowest mass halos \citep[e.g.][]{Artale2015}. However, the uncertainties in the frequency and delay time distributions for when these events should occur, and their relative rarity, preclude including these effects in this \Dissertation.

\subsubsection{SNFeedback}
\label{intro:sec:SNe}

In the simplest model, SNfeedback is treated as the injection of pure thermal energy (and mass) into the fluid element(s) that are host / closest to the chosen site of a SNexplosion. In self-consistent models of star formation and feedback these sites are usually actual star particles, but this does not have to be the case. The first models which included SNfeedback lacked the spatial / mass resolution to resolve individual (or even collective) SN events. This is the source of the commonly known ``overcooling" problem, whereby too little energy is injected into too much mass such that the affected region only reaches modest temperatures ($T \lesssim 10^5$~K, as opposed to $T > 10^7$~K, or so). This temperature is right around the peak of the cooling curve, and thus the energy is rapidly radiated away before it has time to make a significant dynamical impact of the galaxy. This is still a problem today in large-box simulations of cosmological volumes which have insufficient resolution to resolve individual SNe. While some of the initial, ad-hoc fixes for this problem (e.g. turning off cooling for some time in SN affected fluid elements or decoupling SN affected SPH particles from the hydrodynamics for some time) are still in use today, some notable advancements have been made to employ some combination of kinetic, momentum, and thermal energy feedback to attempt to capture SNe in a more consistent fashion \citep[e.g.][]{FIRE,Simpson2015,Hopkins2018}.

High resolution simulations, like those developed in this \Dissertation, can typically rely on using pure thermal energy injection while still avoiding artificial overcooling.\footnote{This is advantageous because it is often substantially easier to implement than kinetic / momentum injection methods while being less susceptible to numerical artifacts, especially for grid-based codes.} This is true for simulations with typical spatial / mass resolutions better than a few pc / 10 M$_{\odot}$ \citep{Simpson2015,Hu2018,Smith2018b}. However, a spatial resolution requirement is a density dependent statement; in this case SNe that occur in the densest gas ($n > 10^{2-3}$~cm$^{-3}$) may still be under-resolved and ineffective. In this case, accounting for feedback processes from massive stars that occur before the first core collapse SN in a given star formation event becomes critical \citep{Hu2016}. Including these physics, as discussed below, greatly reduces the typical ISM densities in which SNe occur, greatly increasing the likelihood that these events are well resolved.

@@

%\setcounter{section}{0}%
%\renewcommand\thesection{\thechapter.\Alph{section}}
%
%\section{Other Works}
%For the sake of maintaining a well-focused \Dissertation -- on stellar feedback and galactic chemical evolution -- we leave out discussion of work done that, while related to galactic evolution,
